Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)

### 2016-03-14 manuscript/supplementary.Rnw
More edits

\section*{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by  tussock grass; and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.

@@

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig. S5-7). OTCs passively increased average ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

@@

The occurrence and severity of fire was initially determined immediately after the 2003 fires\cite{Williams:2006tl}. The severity of the fires in each burnt site was estimated using mean twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}. These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals. Across the 30 burnt sites, the average site twig diameter across ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity. This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.

@@

\subsection{\textit{Modelling shrub seedling growth rates.}}
We modelled the observed height of individual $i$ at May census $t$ as a random realization from a normal distribution truncated at zero:

@@

where $\widehat{height}_{i,t}$ is the estimated maximum height for each individual $i$ at each May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:

@@

We used a logistic equation because it can simulate sigmoidal growth. That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when adult height is reached. We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contains biologically interpretable parameters. Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height that species $j$, to which individual $i$ belongs, can achieve. We did not use the model to estimate $Hmax$ and instead used species specific constants. We did this because our data spanned only the first five years of growth and thus individuals did not reach their maximum heights. We estimated these constants from maximum height data from 5 to 22 long-unburnt sites (depending on the species) to estimated the average maximum height to be approximately \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} and \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively. $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e. the first May measurement). $Days_t$ is the number of days from the beginning of the experiment to each May census $t$.  The experiment ran for a total of \Sexpr{n_days(chamber_dates, FALSE)} days (or approximately \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years). We divided $Days_t$ by 365.25 in order to estimate annual growth rates. The parameter of interest, $R_i$, is the mean annual growth rate for each individual $i$. This parameter controls the slope at inflection (i.e. the maximum annual relative growth rate). When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point. We modelled $R_i$ separately for seedlings growing in bare ground (Equation \ref{eq:r_bg}) and those growing in inter-tussock gaps (Equation \ref{eq:r_tussock}).

@@

\begin{equation}\label{eq:r_bg} 
\begin{aligned}
R_i &=  \alpha_{j[i]} + \beta_{j[i]} \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\alpha_j &\sim \mathcal{N}(\mu_{spp}^{CTL}, \sigma_{spp}^{CTL}) \\
\beta_j &\sim \mathcal{N}(\mu_{spp}^{OTC}, \sigma_{spp}^{OTC}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs}).
\end{aligned}
\end{equation}

Here, the intercept term, $\alpha_{j[i]}$, represents the expected control plot growth rate of species $j$ to which individual $i$ belongs. It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$ which defines the variation between species\textquotesingle growth rates at control plots. The parameter $\beta_{j[i]}$ is the OTC effect for species $j$ to which individual $i$ belongs. $\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the expected OTC growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species\textquotesingle growth rate within OTCs. The terms $\epsilon_{plot_{k[i]}}$ and $\epsilon_i$ represent residual variation at the plot level (for each plot, $k$, within which individual $i$ is located), and the individual level (for each individual $i$), respectively. These error terms were drawn from normal distributions centered on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{obs}$, respectively.

@@

\begin{equation}\label{eq:r_tussock} 
\begin{aligned}
R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs}).
\end{aligned}
\end{equation}

@@

We used Bayesian inference and hence needed to specify prior distributions for all model parameters. In both growth models standard deviation parameters (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25. We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}. In both sets of models the intercept terms ($\mu_{spp}^{CTL}$, $\alpha$), coefficients (i.e. inter-tusscok model: $\beta_1$, $\beta_2$ and $\beta_3$) and hyperparameter (i.e. in the multispecies model) for the mean OTC effect, $\mu_{spp}^{OTC}$, were estimated from normal priors centered on zero with a standard deviation of 0.5. We chose this prior as it spanned the range of likely growth trajectories from very slow to very fast growing.

\subsection{\textit{Modelling shrub seedling mortality rates.}}
We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realizations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

@@

As with growth rates, we modelled each individual's cumulative hazard, \\
$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches and those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the time lapsed (in years) between censuses, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.

@@

Here, the intercept term, $\alpha_{j[i]}$, represents the expected instantaneous hazard rate in a control plot for species $j$ to which individual $i$ belongs. It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species\textquotesingle instantaneous hazard rates in control plots. The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs. $\beta_{j[i]}$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the OTC expected probability of death for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$, which defines the variation between species’ instantaneous hazard rates in OTCs. The parameters $\epsilon_{plot_k}$ and $\epsilon_i$ are residuals at the plot and individual level, respectively.

@@

\subsection{\textit{Modelling Grevillea recruitment rates.}}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realization from a negative binomial distribution:

@@

where the intercept, $\alpha$, represents the average log recruitment in control plots in 2011. The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in the log number of recruits from 2011 to 2012. $\epsilon_{plot_k}$ is the residual error at plot $k$.

@@

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. \textgreater 25 cm), these data were incomplete. We estimated these censored data using a censored model (see below). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

\subsection{\textit{Modelling censored Poa distances.}}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ cardinal distance data used to calculate the average inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. \textgreater  25 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.

Specifically, we first modelled the observed (non-censored) cardinal distances as random realizations from a lognormal distribution:

@@

where $\textrm{ln}(\mu_{i,t})$ is the log estimated median radius for inter-tussock gap, at census $t$, and $\sigma_{obs}$ is the standard deviation of the observation error.

@@

\subsection{\textit{Modelling drivers of seedling density.}}
We examined drivers of seedling density by modelling the number of seedlings square metre plot $i$ as a random realization from a Poisson distribution:

@@

\subsection{\textit{Modelling drivers of seedling maximum heights.}}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realizations from a log-normal distribution:

@@

Here, we estimated $\mu_{site}$ and coefficients ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5. These priors encapsulated our prior knowledge that seedling maximum heights would be less than 150 cm. All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.

\subsection{\textit{Data inputs into density and maximum height models.}}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity $Severity_j$, was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} < 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}. 
\pagebreak
\section*{Supplementary figures}
\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs1}
\end{minipage}
\end{suppfigure}
\pagebreak

@@

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS3.pdf}
\caption{\textbf{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.} Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively. orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.} 
\label{figs3}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS4.pdf}
\caption{\textbf{Mean maximum seedling height along fire severity (minimum twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. (\textbf{a} \& \textbf{b}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{c} \& \textbf{d}) fire severity, (\textbf{e} \& \textbf{f}) altitude and (\textbf{g} \& \textbf{h}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs4}
\end{minipage}
\end{suppfigure}
\pagebreak

>>>>

\section*{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.

@@

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions\cite{IPCC:2013vv} (Supplementary Fig. S5-7). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsection{\textit{Data inputs for growth, mortality and recruitment models.}}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations\cite{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

The occurrence and severity of fire was determined immediately after the 2003 fires\cite{Williams:2006tl}. The severity of the fires in each burnt site was estimated using mean twig diameter, a proxy measure of fire severity\cite{Whight:1999vw}. These twig diameters were recorded between 1 and 10 months post-fire from 10 replicate \textit{Grevillea australis} individuals per site. Across the 30 burnt sites, the average site twig diameter ranged from 2 to 13 mm, with smaller diameters assumed to correspond to lower fire intensity. This encompassed the full range of fire severity from lightly scorched sites containing shrubs with only the tips of the outer branches burnt, to severely burnt sites where the majority of large shrub branches were consumed.

\subsection{\textit{Data inputs into density and maximum height models.}}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations\cite{Gelman:2007te}.

@@

\subsection{\textit{Modelling shrub seedling growth rates.}}
We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:

@@

where $\widehat{height}_{i,t}$ is the estimated maximum height for individual $i$ at May census $t$ and $\sigma_{obs}$ is the standard deviation of the observation error. The parameter $\widehat{height}_{i,t}$ was estimated using a logistic parameterisation of the unified Richards equation\cite{Tjorve:2010bk}:

@@

We used a logistic equation because it can simulate sigmoidal growth. That is, a curve whereby initial seedling growth is slow before increasing in speed and finally levelling off when maximum adult height is reached. We used this particular parameterisation because it accounts for seedling size at the beginning of an experiment and contains biologically interpretable parameters. Here, $Hmax_{j[i]}$ is the asymptotic height: the average mature height that species $j$, to which individual $i$ belongs, can achieve. We did not use the model to estimate $Hmax$ and instead used species specific constants. We did this because our data spanned only the first five years of growth and thus individuals did not reach their maximum heights. We estimated these constants from maximum height data from 5 to 22 long-unburnt sites (depending on the species) to estimated the average maximum height to be approximately \Sexpr{hmax_values(hmax,'Asterolasia')}, \Sexpr{hmax_values(hmax,'Grevillea')}, \Sexpr{hmax_values(hmax,'Phebalium')} and \Sexpr{hmax_values(hmax,'Prostanthera')} cm for \textit{Asterolasia}, \textit{Grevillea}, \textit{Phebalium} and \textit{Prostanthera}, respectively. $Hinit_i$ is the height of individual $i$ at the start of the experiment (i.e. the first May measurement). $Days_t$ is the number of days from the beginning of the experiment to each May census $t$.  The experiment ran for a total of \Sexpr{n_days(chamber_dates, FALSE)} days (or approximately \Sexpr{round(n_days(chamber_dates, FALSE)/365.25)} years). We divided $Days_t$ by 365.25 in order to estimate annual growth rates. The parameter of interest, $R_i$, is the mean annual growth rate for each individual $i$. This parameter controls the slope at inflection (i.e. the maximum annual relative growth rate). When multiplied by the asymptotic height, $Hmax$, it yields the maximum annual growth rate at the inflection point. We modelled $R_i$ separately for seedlings growing in bare ground (Equation \ref{eq:r_bg}) and those growing in inter-tussock gaps (Equation \ref{eq:r_tussock}).

@@

\begin{equation}\label{eq:r_bg} 
\begin{aligned}
R_i &=  \alpha_{j[i]} + \beta_{j[i]} \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\alpha_j &\sim \mathcal{N}(\mu_{spp}^{CTL}, \sigma_{spp}^{CTL}) \\
\beta_j &\sim \mathcal{N}(\mu_{spp}^{OTC}, \sigma_{spp}^{OTC}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

Here, the intercept term, $\alpha_{j[i]}$, represents the expected control plot growth rate of species $j$ to which individual $i$ belongs. It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$ (the expected control plot growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$ which defines the variation between species\textquotesingle{} growth rates at control plots. The parameter $\beta_{j[i]}$ is the OTC effect for species $j$ to which individual $i$ belongs. $\beta_j$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the expected OTC growth rate for an average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$ which defines the variation between species\textquotesingle{} growth rate within OTCs. The terms $\epsilon_{plot_{k[i]}}$ and $\epsilon_i$ represent residual variation at the plot level (for each plot, $k$, within which individual $i$ is located), and the individual level (for each individual $i$), respectively. These error terms were drawn from normal distributions centered on zero with standard deviations that defined the variation amongst plots, $\sigma_{plot}$, or amongst individuals, $\sigma_{ind}$, respectively.

@@

\begin{equation}\label{eq:r_tussock} 
\begin{aligned}
R_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

We used Bayesian inference and hence needed to specify prior distributions for all model parameters. In both growth models standard deviation parameters (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$, $\sigma_{ind}$ and $\sigma_{obs}$) were all estimated from weakly informative positive half Cauchy priors with scale parameters of 25. We used these Cauchy priors in order to prevent over-inflated estimates of variances due to the small number of plots and species\cite{Gelman:2006uw}. In both sets of models the intercept terms ($\mu_{spp}^{CTL}$, $\alpha$), coefficients (i.e. inter-tusscok model: $\beta_1$, $\beta_2$ and $\beta_3$) and hyperparameter (i.e. in the multispecies model) for the mean OTC effect, $\mu_{spp}^{OTC}$, were estimated from normal priors centered on zero with a standard deviation of 0.5. We chose this prior as it spanned the range of likely growth trajectories from very slow to very fast growing.

\subsection{\textit{Modelling shrub seedling mortality rates.}}
We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realisations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

@@

As with growth rates, we modelled each individual's cumulative hazard, \\
$Cumulative\;hazard_{i,t}$, separately for seedlings growing in bare ground patches and those in inter-tussock gaps. In both cases, $Cumulative\;hazard_{i,t}$, was modelled on the complementary log-log scale as a function of an instantaneous hazard rate for each individual, $\lambda_i$, multiplied by the census interval (in years) between censuses, $t$. By doing this our regression model becomes equivalent to a parametric survival analysis\cite{Kleinbaum:2012tn}.

@@

Here, the intercept term, $\alpha_{j[i]}$, represents the expected instantaneous hazard rate in a control plot for species $j$ to which individual $i$ belongs. It is estimated from a normal distribution centered on $\mu_{spp}^{CTL}$  (the control plot expected instantaneous hazard rate for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{CTL}$, which defines the variation between species\textquotesingle{} instantaneous hazard rates in control plots. The parameter $\beta_{j[i]}$ is the OTC effect for each species $j$ to which individual $i$ belongs. $\beta_{j[i]}$ is also estimated from a normal distribution, but unlike $\alpha_j$, it is centered on $\mu_{spp}^{OTC}$, (the OTC expected probability of death for the average hypothetical species) with a standard deviation, $\sigma_{spp}^{OTC}$, which defines the variation between species’ instantaneous hazard rates in OTCs. The parameters $\epsilon_{plot_k}$ and $\epsilon_i$ are residuals at the plot and individual level, respectively.

@@

\subsection{\textit{Modelling Grevillea recruitment rates.}}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:

@@

where the intercept, $\alpha$, represents the average log recruitment in control plots in 2011. The parameter $\beta_1$ is the effects of OTC and $\beta_2$ describes the change in the log number of recruits from 2011 (i.e. $Census$ = 0) to 2012 (i.e. $Census$ = 1). $\epsilon_{plot_k}$ is the residual error at plot $k$.

@@

\subsection{\textit{Modelling censored Poa distances.}}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. $\geq$  26 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.

Specifically, we first modelled the observed (non-censored) cardinal distances as random realisations from a lognormal distribution:

@@

where $\textrm{ln}(\mu_{i,t})$ is the log estimated median radius for inter-tussock gap $i$, at census $t$, and $\sigma_{obs}$ is the standard deviation of the observation error.

@@

\subsection{\textit{Modelling drivers of seedling density.}}
We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:

@@

\subsection{\textit{Modelling drivers of seedling maximum heights.}}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation (equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C\cite{Brown:1989vx}) influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:

@@

Here, we estimated $\mu_{site}$ and coefficients ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5. These priors encapsulated our prior knowledge that seedling maximum heights would be less than 150 cm. All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.
\pagebreak
\section*{Supplementary figures}
\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs1}
\end{minipage}
\end{suppfigure}
\pagebreak

@@

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS3.pdf}
\caption{\textbf{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs.} Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively. Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.} 
\label{figs3}
\end{minipage}
\end{suppfigure}
\pagebreak

\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS4.pdf}
\caption{\textbf{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires.} Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. (\textbf{a} \& \textbf{b}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{c} \& \textbf{d}) fire severity, (\textbf{e} \& \textbf{f}) altitude and (\textbf{g} \& \textbf{h}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs4}
\end{minipage}
\end{suppfigure}
\pagebreak
